### 2024-02-15 outputs/paper/paper.qmd
added knitr

---
title: "Voting Patterns and Socio-Economic: A Twenty-Year Analysis of Brazil's Developmental Dynamics"
author: "Hao Wang / Siqi Fei / Xuecheng Gao"
thanks: "Code and data are available at: https://github.com/groundUofT888/Term-Paper-2.git ; Replication on Social Science Reproduction platform available at: https://doi.org/10.48152/ssrp-qxvj-fz21"
date: "February 12, 2024"
date-format: long
format: 
  pdf: 
    include-in-header: 
      text: |
        \usepackage{lscape}
        \newcommand{\blandscape}{\begin{landscape}}
        \newcommand{\elandscape}{\end{landscape}}
    documentclass: article
    geometry: margin = 1in
    abstract: " "
    number-sections: true
output: 
  bookdown::pdf_document2
toc: false
fig_caption: yes 
nocite: '@*'
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---
\newpage
```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# install.packages("readr")
# install.packages("here")
# install.packages("dplyr")
# install.packages("tidyverse")
# install.packages("ggplot2")
# install.packages("knitr")

library(here)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(knitr)

data_path <- here("outputs", "edited_data", "election_cleaned.csv")
election <- read.csv(data_path)
data_path2 <- here("outputs", "edited_data", "schools.csv")
school <- read.csv(data_path2)
```

```{=tex}
\begin{center}
\textbf{\large Abstract}
\end{center}
```
```{=tex}
\begin{quote}
The paper replicates data to analyze economic indicators such as income per capita, the low-income rate, and the unemployment rate during two-round elections in Brazil, while considering the dropout rate as a main educational indicator and exploring the impact of electoral provisions on citizens, the country's economy, and society. We conduct secondary research on Brazil’s economic and education data and find that politicians’ campaign strategies influence voters in two-round electoral systems, further affecting the local economy and policy implementation.

We hope that by analyzing the Brazilian electoral system, we can provide new perspectives and ideas for studying the impact of the world's political electoral systems. We call for future investigations in this area to provide more information, and we also look forward to offering useful suggestions and inspirations for Brazil's social development and economic construction.
\end{quote}
```

# Introduction

In 2022, Brazil's presidential election broke new grounds with an
unprecedented turnout of over 150 million voters, capturing the
attention of national and international media. The world had seen not
just about the political tactics of the candidates, but also the
public's vested interest in how these elections could shape their
day-to-day lives. This paper aims to study the influence of Brazil's
electoral system on three critical socio-economic indicators: income
level, unemployment rate and educational resources.With a comprehensive
look at recent data and trends, we will investigate how the details of
voting and representation create a chain reaction throughout the
economy, affecting everything from job opportunities to the quality of
education. By collecting together statistical analysis with real-world
implications, we'll provide a clearer picture of the electoral system's
impact on the Brazilian populace and discuss the broader implications of
these findings on social equality and economic stability.

# Background

## Two-round elections in Brazil

Brazil's two-round electoral system is a relatively common electoral
method. The mode of election is to select the winner after two votings.
In the first-round voters can choose the candidate they support among
multiple candidates. If one of the candidates gets more than half of the
votes in the first round, then he can be directly elected. If no
candidate receives more than half of the votes in the first round, a
second round will take place. In the second round of voting, voters will
choose between the two candidates who received the highest votes in the
first round. Election regulations：1. Voters must be registered at least
151 days before the election. 2. Voting is compulsory for all literate
Pakistani citizens between the ages of 18 and 69. 3. The election is
held on the first Sunday in October.

## Economic and social issues in Brazil

Major social and economic issues in Brazil. First, Brazil’s poverty gap.
There are many depressed areas in Brazil. Many people live below the
poverty line and lack basic living conditions such as housing,
education, medical care, and jobs. This leads to social instability.
Second, the high crime rate. Some cities in Brazil have high crime rates
and violence problems, such as robberies, gun proliferation, and drug
trafficking. These problems mainly stem from poverty caused by uneven
distribution of wealth and the deterioration of local security. Third,
Brazil has high unemployment. The number of unemployed people is
increasing in Brazil, and many structural unemployment results in high
unemployment rates for a large number of low-skilled laborers. Also
lower wages for many people worsen poverty and inequality. Fourth, there
are issues with the education system. Brazil's education system has many
problems, such as uneven education quality, uneven allocation of
government funds for education, shortage of professors, and high dropout
rates.

# Data

## Source

This paper replicates the survey data that was originally collected by
Moya Chin, AMERICAN ECONOMIC JOURNAL: APPLIED ECONOMICS VOL. 15, NO. 3,
JULY 2023 (pp. 183-209).

It uses data sets constructed from four data sources: - 1) Brazil
municipal election data (Tribunal Superior Eleitoral, 1996-2016); - 2)
Brazil demographic censuses (Institute Brasileiro de Geografia e
Estatística, 1980, 1991, 2000, 2010); - 3) Brazil school census
(Ministério da Educação, 1997-2016); - 4) nighttime lights (National
Oceanic and Atmospheric Administration, 1997-2016a,-)

R[@citeR] is the language and environment used for this
analysis,supported by the robust features of Posit Cloud as our primary
integrated development environment (IDE). Library includes
`ggplot2`[@ggplot2], `here`[@here].

Git(version 2.39.2) is used as version control.

## Features

### Data Sets Overview

Our comprehensive data collection provides a longitudinal perspective on
the socio-economic and political landscape of Brazil, spanning from 1996
to 2016. This period covers six Brazilian electoral cycles, offering a
unique vantage point for analyzing socio-economic fluctuations and
trends through detailed indicators.

### 1. Election Dataset (election_cleaned.csv)

-   Path: outputs/edited_data/election_cleaned.csv

-   Description: Contains socio-economic indicators for Brazil,
    capturing fluctuations and trends across six electoral cycles from
    1996 to 2016.
    ```{r echo = FALSE, results = 'asis'}
    election_vars <- tibble(
      Variable = c("cut_dist", "elect_year", "inc_post", "inc_0_50_post", "edu_univ", "unempl_post", "tse_code"),
      Description = c("Geographical identifier for districts.",
                      "Year of the election.",
                      "Per capita income post-election.",
                      "Income for the bottom 50% post-election.",
                      "Educational attainment/resources.",
                      "Post-election unemployment rates.",
                      "Electoral zones/districts identifier.")
    )

    kable(election_vars, format = "markdown", caption = "Election Dataset Variables")

    ```
\newpage
    ### 2. Schools Dataset (schools.csv)

-   Path: outputs/edited_data/schools.csv

-   Description: Offers insights into Brazil's socio-economic and political landscape through educational and economic measures.
```{r echo = FALSE, results ='asis'}

# Schools Dataset Variables
schools_vars <- tibble(
  Variable = c("ibge7_code", "cut_dist", "pop_den_pre", "dep_pca_ptile_mean", "MAT_DROP_mean"),
  Description = c("Geographic identifier (7-digit IBGE code).",
                  "Geographical identifier for districts.",
                  "Pre-election population density.",
                  "Mean dependency ratio across percentiles.",
                  "Mean drop-out rate in schools.")
)

kable(schools_vars, format = "markdown", caption = "Schools Dataset Variables")

```
\renewcommand{\arraystretch}{2}

\newpage

# Result

The sample size, comprising 33,188 observations, was determined to ensure representativeness and statistical validity. The dataset encompasses 5,567 data points per election year, reflecting a robust cross-section of Brazil's population and socio-economic conditions during the study period.

## Income per Capita

```{r}
#| label: fig-terc
#| fig-cap: "Income per capita"
#| echo: false
#| message: false
#| warning: false
#| fig.width: 3
#| fig.height: 2

p <- ggplot(data = election, aes(x = tse_code, y = inc_post)) +
  geom_point(color = "blue") +  # Scatter plot points
  labs(x = "Number of registered voters", y = "Income per capita")
p + geom_smooth(method = "lm", color = "red")
```

For the above [@fig-terc],from 1996 to 2016, the trend line indicates a
slight positive correlation between the number of registered voters and
income per capita. This could suggest that in areas with more registered
voters, there may be a slightly higher income per capita.

## Low Income Rate

```{r}
#| label: fig-excess
#| fig-cap: "Low income rate"
#| echo: false
#| message: false
#| warning: false
#| fig.width: 3
#| fig.height: 2

cutoff_value <-60000
election$cutoff_side <- ifelse(election$tse_code < cutoff_value, "Below", "Above")

p <- ggplot(data = election, aes(x = tse_code, y = inc_0_50_post)) +
  geom_point(color = "blue") +  # Scatter plot points
  labs(x = "Number of registered voters", y = "Low income rate") +
  geom_vline(xintercept = cutoff_value, linetype = "dashed", color = "black") + # Cutoff line
  geom_smooth(data = subset(election, tse_code < cutoff_value), 
              aes(x = tse_code, y = inc_0_50_post), 
              method = "lm", color = "red") + 
  geom_smooth(data = subset(election, tse_code >= cutoff_value), 
              aes(x = tse_code, y = inc_0_50_post), 
              method = "lm", color = "yellow") 
p

```

We defined the lowest 50% income as low income. The plot [@fig-excess]
appears to show a slight negative correlation, as indicated by the
downward trend line. This suggests that as the number of registered
voters increases, the low-income rate decreases slightly. However, the
wide spread of the data points indicates that many other factors likely
influence the low-income rate beyond just the number of registered
voters. \newpage

## Unemployment rate

```{r}
#| label: fig-ecoil
#| fig-cap: Unemployment rate
#| echo: false
#| message: false
#| warning: false
#| fig.width: 3
#| fig.height: 2

p <- ggplot(data = election, aes(x = tse_code, y = unempl_post)) +
  geom_point(color = "blue") +  # Scatter plot points
  labs(x = "Number of registered voters", y = "Unemployment rate")
p + geom_smooth(method = "lm", color = "red")
```

Analyzing the [@fig-ecoil], there are noticeable differences. The plot
shows a vast spread of unemployment rates across varying numbers of
registered voters, with a trend line that remains relatively flat. This
suggests there isn't a strong or clear correlation between the number of
registered voters and the unemployment rate within the represented data
set.

## Educational Resources

```{r}
#| label: fig-model
#| fig-cap: "education"
#| echo: false
#| message: false
#| warning: false
#| fig.width: 3
#| fig.height: 2

cutoff_value <-60000
school$cutoff_side <- ifelse(school$tse_code < cutoff_value, "Below", "Above")

p <- ggplot(data = school, aes(x = tse_code, y = MAT_DROP_mean)) +
  geom_point(color = "blue") +  
  labs(x = "Number of registered voters", y = "Drop-our rate") +
  geom_vline(xintercept = cutoff_value, linetype = "dashed", color = "black") + # Cutoff line
  geom_smooth(data = subset(election, tse_code < cutoff_value), 
              aes(x = tse_code, y = inc_0_50_post), 
              method = "lm", color = "red") + 
  geom_smooth(data = subset(election, tse_code >= cutoff_value), 
              aes(x = tse_code, y = inc_0_50_post), 
              method = "lm", color = "red") 
p 
```

Analyzing the plot presented in [@fig-model], titled "Educational
Resources," we see a substantial variation in dropout rates across a
range of registered voter numbers. The dual trend lines, both
descending, suggest that areas with more registered voters tend to have
lower dropout rates, yet the change in the slope at around the 50,000
voter mark could imply different dynamics in larger districts. Despite
the overall negative correlation, the broad spread of the data points,
especially in regions with fewer voters, indicates that voter numbers
alone may not be a definitive predictor of dropout rates. The
relationship is not uniform across the dataset, highlighting the
complexity of factors that influence educational outcomes. \newpage

## Overview

```{r}
#| label: fig-hist
#| fig-cap: "overview histogram"
#| echo: false
#| message: false
#| warning: false
#| fig.width: 8
#| fig.height: 7

data_long <- pivot_longer(election, cols = c("inc_0_50_post", "inc_post", "unempl_post"), 
                          names_to = "Variable", values_to = "Value") %>%
  mutate(DescriptiveLabel = case_when(
    Variable == "inc_0_50_post" ~ "Low Income",
    Variable == "inc_post" ~ "Income per Capita",
    Variable == "unempl_post" ~ "Unemployment Rate"
  ))

ggplot(data_long, aes(x=Value, fill=Variable)) + 
  geom_histogram(position="dodge", binwidth=1) + 
  facet_wrap(~ DescriptiveLabel, scales = "free_x") + 
  theme_minimal() + 
  ggtitle("Distribution of Economic Indicators") +
  xlab("Value") +
  ylab("Frequency") +
  scale_fill_manual(values = c("inc_0_50_post" = "blue", "inc_post" = "red", "unempl_post" = "green"),
  labels = c("Low Income Rate", "Income per Capita", "Unemployment Rate")
)

```

Analyzing the histogram in [@fig-hist], we see distinct distributions
for three key economic indicators. The Income per Capita is broadly
spread, suggesting varied economic conditions. In contrast, the Low
Income Rate displays a bimodal distribution, hinting at two prevalent
income groups within the population. The Unemployment Rate is
predominantly concentrated around a lower value but with a long tail
toward higher rates, indicating that most of the population experiences
lower unemployment, with fewer instances of high unemployment. This
visual representation underscores the complexity of economic health and
inequality within the dataset. \renewcommand{\arraystretch}{1.5}

\newpage

```{=tex}
\begin{landscape}
```
```{=tex}
\end{landscape}
```
```{r, message=FALSE, message=FALSE, echo=FALSE, include=FALSE}
#| label: draft code (not used)

```

# Discussion

## Findings

Our examination of the data presented by Moya Chin through a regression discontinuity design yields nuanced insights into the effects of Brazil's electoral system on socio-economic variables. We delineate two primary findings from our analysis:

* Electoral Influence on Economic Conditions: Contrary to the initial reading of the data, the two-round elections in Brazil do not show a definitive positive impact on the broader downstream economy within the period studied. While there is an observable increase in per capita income and a decrease in the number of low-income individuals during election periods, these trends are not statistically significant when considering the larger socio-economic context. The data suggests that any changes in income levels and unemployment rates associated with the electoral process are subtle and may require a more extended period to be fully realized and understood.

* Campaign Strategies and Voter Demographics: Our findings align with the hypothesis that candidates during elections strategically target demographics that are most likely to influence the outcome, specifically low-income individuals and those with lower educational attainment. This is reflected in the allocation of resources, such as improvements in school infrastructure and public education facilities, which are more likely to directly benefit these groups. While this approach does translate into immediate educational gains, such as reduced dropout rates and increased literacy rates, it is less clear how these short-term improvements translate into long-term socio-economic development.

From Moya Chin's work, it is evident that the behavior of politicians during election campaigns has a tangible impact on the distribution of public resources, which can shift dramatically post-election. While two-round elections may encourage candidates to commit more resources to public education and economic support, this study finds that the long-term economic effects, such as sustained increases in income or reductions in unemployment, are not as immediate or discernible within the data.

Additionally, the presence of political tactics aimed at low-income voters suggests a complex interplay between electoral strategies and voter demographics. Despite the potential short-term benefits to public education and economic assistance that may arise from two-round elections, our analysis indicates that addressing Brazil's economic and social challenges will likely require a more extended and multifaceted approach.

As the conclusion, while the electoral system does have some influence on socio-economic outcomes, the extent and durability of these effects are contingent on a variety of factors and warrant further, longitudinal study to fully comprehend their scope and longevity.

## Ethical Implications

The datasets and original sources used in their paper come from Brazil
demographic censuses and Brazil school census. Demography refers to
summarizing the demographic characteristics of a community, such as
gender, age, race, education level, socioeconomic status, etc. Ethical
issues may arise if used improperly during the collection, analysis,
interpretation, and dissemination of demographic data.

Demographics impact minority communities, which can lead to structural
inequalities. When analyzing demographic variables, the author selected
specific variables for analysis based on his research questions and
hypotheses. However, these highly characteristic data will have a
negative impact on some social groups. For example, high income, low
income, dropout rate, etc. This may reinforce social hierarchies in
which social groups are compared, exacerbating prejudice and stereotypes
and causing harm to minority populations. In the paper, the author
compares the interest of high-income people and low-income people in
voting. This may cause certain social conflicts and encourage the idea
of inequality among different groups. When explaining differences
between social groups in the conclusion of the author's paper, it can
lead to increased social conflict.

## Accounting for Bias

In the paper, Moya Chin proposes a stylized model that explores the
motivations of politicians to appeal to a wider range of voters in
elections and to provide public goods in different ways. The author uses
this model to observe the Economic Consequences of Electoral Rules in
Brazil. The data used by the author mainly comes from Brazil demographic
censuses, Brazil school Census, and Brazil municipal election data. Bias
can occur when collecting quantitative and qualitative data.

The problem with these data sets and original sources is that the
variables used by the authors to construct the equipment index come from
the Brazilian School Census and the Brazilian demographic censuses.
Considering Brazil's national conditions and the government's influence
on the region, Response Bias may exist in the data set of Brazil
demographic censuses. Census respondents may choose not to participate
in the census or provide inaccurate information because of privacy
concerns or distrust of the government. The school census data comes
from students in schools. Considering the lack of educational resources
in poor areas of Brazil, there will be a sampling frame bias. The
earliest census data used by the author is from 1980. At that time, in
some remote or poor areas, it may be more difficult for enumerators to
reach or collect data due to lack of infrastructure and information
technology. This may lead to underestimation or inaccurate population
data in these areas, and the sample may not be representative of the
characteristics of the entire population.

## Limitations

Our analysis is based on a data set from the Moya Chin literature, so
most of our analysis is limited to the original survey situation.

Because the data are not comprehensive enough, we cannot compare with
many variables in the original data, such as whether political
intervention and media propaganda affect voters' thoughts and voting
results. Excessive propaganda of policies and political interference are
suspected of manipulating voters and leading to distorted election
results. Also the political contributions a politician receives during
an election can influence the outcome of the election.

The model created by Moya Chin cannot be fully applied during our
analysis, which results in our analysis not being as complete as Moya
Chin described. So we have to give up some variables.

## Future Research

We believe that this research direction can receive more investment in
the future. We hope that the relationship between electoral systems,
education, and socioeconomic outcomes can be further explored in the
future. We hope to observe and predict the impacts of elections over the
long term to fully understand the impact of the electoral system on
economic indicators, and to reveal the possible long-term effects of
changes in the electoral system.

\clearpage

# References {.unnumbered}